Discover governing reactions from concentration data

### 2017-10-27 docs/manuscript.tex
manuscript: wip tex

% \usepackage{graphicx}

@@

	\section*{The method}
	The underlying model that we want to fit the data to is a law of mass action type dynamical system. To this end, let $S$ be number of species and $T$ be the number observations. The concentration data at a time $t\in [1, T]$ can be represented by a vector
	\[
	\mathbf{x}(t)=\begin{pmatrix}
	x_1(t)\\ \vdots \\ x_S(t)
	\end{pmatrix}\in \mathbb{R}^S.
	\]
	Further, one can choose $R$ possible ansatz reactions with their respective reaction function
	\[
	\theta_r(\textbf{x}(t))=\begin{pmatrix}
	\theta_{r,1}(\textbf{x}(t)) \\ \vdots \\ \theta_{r,S}(\textbf{x}(t))
	\end{pmatrix}
	\]
	so that the change of concentration for species $i$ at time $t$ is represented by the dynamical system
	\[
	\dot{\textbf{x}}_i(t) = \sum_{r=1}^{R}\theta_{r,i}(\textbf{x}(t))\xi_r,
	\]
	where $\xi_r$ are the to-be estimated macroscopic reaction rates.

@@

	\newpage
	\bibliographystyle{alpha}
	\bibliography{bibliography.bib}

>>>>

% \usepackage{graphicx}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

	\section*{The method}
	The underlying model that we want to fit the data to is a law of mass action type dynamical system. To this end, let $S$ be the number of species, then the concentration data at a time $t$ can be represented by a vector
	\[
	\mathbf{x}(t)=\begin{pmatrix}
	x_1(t)\\ \vdots \\ x_S(t)
	\end{pmatrix}\in \mathbb{R}^S.
	\]
	Further, one can choose $R$ possible ansatz reactions with their respective reaction function
	\[
	\textbf{y}_r(\textbf{x}(t))=\begin{pmatrix}
	y_{r,1}(\textbf{x}(t)) \\ \vdots \\ y_{r,S}(\textbf{x}(t))
	\end{pmatrix}
	\]
	so that the change of concentration for species $i$ at time $t_i$, is represented by the dynamical system
	\[
	\dot{\textbf{x}}_i(t) = \sum_{r=1}^{R}y_{r,i}(\textbf{x}(t))\xi_r,\quad i=1,\ldots, S,
	\]
	where $\xi_r$ are the to-be estimated macroscopic rates.

	When presented with a time series with $T$ observations, the data can be represented as a matrix
	\[
	\textbf{X} = \begin{pmatrix}
		x_1(t_1) & x_2(t_1) & \cdots & x_S(t_1) \\
		x_1(t_2) & x_2(t_2) & \cdots & x_S(t_2) \\
		\vdots   & \vdots   & \ddots & \vdots   \\
		x_1(t_T) & x_2(t_T) & \cdots & x_S(t_T)
	\end{pmatrix} \in \mathbb{R}^{T\times S}.
	\]
	Given this matrix, one can propose a library $\Theta(\textbf{X}) = \begin{pmatrix} \theta_1(\textbf{X}) & \theta_2(\textbf{X}) & \cdots & \theta_R(\textbf{X}) \end{pmatrix}$ of $R$ ansatz reactions with corresponding reaction functions
	\[
	\theta_r(\textbf{X}) = \begin{pmatrix}
	\textbf{y}_r(\textbf{X}_1)^T \\ \vdots \\ \textbf{y}_r(\textbf{X}_T)^T
	\end{pmatrix}\in \mathbb{R}^{T\times S},\quad r=1,\ldots,R,
	\]
	where $\textbf{X}_i$ denotes the $i$-th row in $X$. Applying the concentration trajectory to the library yields a data tensor $\Theta(\textbf{X})\in\mathbb{R}^{T\times S\times R}$. Following the approach of SINDy, the goal is now to find coefficients $\Xi = \begin{pmatrix} \xi_1 & \xi_2 & \cdots & \xi_R
	\end{pmatrix}^T$, so that
	\[
	\dot{\textbf{X}} = \Theta(\textbf{X})\Xi = \sum_{r=1}^{R}\theta_r(\textbf{X})\xi_r.
	\]
	In particular, the system is linear in the coefficients $\Xi$, which makes sparse regression tools such as LASSO \cite{Tibshirani1996, Hastie2009} applicable. To this end, one can consider the minimization problem to find $\hat{\Xi}$ such that
	\begin{align}
		\hat{\Xi} = \argmin_{\Xi}\left( \frac{1}{2T}\left\| \dot{\textbf{X}} - \Theta(\textbf{X})\Xi \right\|_F^2 + \alpha\|\Xi\|_1 \right) \quad \text{subject to }\Xi \geq 0,
	\label{method:minimizationproblem}\end{align}
	where $\alpha\geq 0$ is a sparsity inducing hyperparameter. For $\alpha=0$ this problem reduces to constrained least-squares. For solving (\ref{method:minimizationproblem}) we applied the sequential least squares minimizer SLSQP, originally described in \cite{Kraft1988}, contained in the software package SciPy \cite{SciPy}.

@@

	\newpage
% 	\bibliographystyle{alpha}
	\bibliographystyle{abbrv}
	\bibliography{bibliography.bib}
